---
title: "Figures for Fullagar et al. 2018"
author:
  - Ben Marwick
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
    bookdown::html_document2:
      fig_caption: yes
      reference_docx: "../templates/template.docx" # Insert path for the DOCX file
bibliography: references.bib
csl: "../templates/journal-of-archaeological-science.csl" # Insert path for the bib-style

---


<!-- This is the format for text comments that will be ignored during renderings. Do not put R code in these comments because it will not be ignored. -->

```{r, setup, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  warning = FALSE,
  message = FALSE,
  echo = TRUE,
  comment = "#>",
  fig.path = "../figures/"
)

# library(dauwrs) # Or use devtools::load_all('.', quiet = T) if your code is in script files, rather than as functions in the `/R` diretory
```

# Making the map

First we make the polygons to highlight regions on the map. This has to be done interactively, by hand. So we just do it once and then leave it. 

```{r demo, eval = FALSE}
library(mapview)
library(mapedit)

# create polygons
western_desert <- 
  mapview() %>%
  editMap()

western_desert_poly <- western_desert$finished

# edit this one some more...

western_desert2 <- 
  mapview(western_desert_poly) %>%
  editMap("western_desert_poly")

western_desert2_poly <- western_desert2$edited

# but why is this necessary?
names(western_desert2_poly)[2]  <- "feature_type"
western_desert2_poly$feature_type <- "polygon"

# next one
queensland_rainforest <- 
  mapview() %>%
  editMap()

queensland_rainforest_poly <- queensland_rainforest$finished

kimberley_region <- 
  mapview() %>%
  editMap()

kimberley_region_poly <- kimberley_region$finished

arnhem_land <- 
  mapview() %>%
  editMap()

arnhem_land_poly <- arnhem_land$finished

# see how they look
roi_polys_rb <-rbind(western_desert2_poly,
                     queensland_rainforest_poly,
                     kimberley_region_poly,
                     arnhem_land_poly)

ggplot(roi_polys_rb) +
  geom_sf()

# save a shapefile
sf::st_write(roi_polys_rb, "../figures/roi_polys_rb.shp")

# save them to disk
saveRDS(roi_polys_rb, "../figures/regions-of-interest-polygons.rds")
```

Now we have a shapefile with the regions as polygons, we can move to the next step of plotting the sites on a basemap. This is reproducible from the code. 

```{r make-the-map}

# read in data ---
site_coords <- readxl::read_excel("../data/raw_data/Fullagar-Hayes-Marwick-grinding-site-locations.xlsx")
region_label_coords <- readxl::read_excel("../data/raw_data/Fullagar-Hayes-Marwick-grinding-site-locations.xlsx", sheet = 2)
region_label_coords$label_name <- stringr::str_remove_all(region_label_coords$label_name, "\r") 

# head(site_coords)

# plot map -----
library(ggmap)
library(ggrepel)
library(legendMap) # devtools::install_github("3wen/legendMap")

### Set a range
mean_lat <- mean(site_coords$lat)                
mean_lon <- mean(site_coords$long)- 10 # bump to make Australia central

### Get a map
base_map <- 
  get_map(location = c(lon =mean_lon, 
                       lat = mean_lat), 
               zoom = 4,
               maptype = "roadmap", 
               source = "google")

# First want to indicate:
# Semi arid/arid regions of Aus 
# Western Desert
# North Queensland rainforest 
# Kimberley region
# Arnhem Land

# Read in the shapefile, then fortify for use with ggplot2
shapefile_df <- fortify(rgdal::readOGR("../figures/roi_polys_rb.shp"))

# add shapefile site map
map_with_sites <- 
  ggmap(base_map) +
  geom_polygon(data = shapefile_df, 
               aes(x = long, 
                   y = lat, 
                   group = group), 
               fill = 'gray60', 
               alpha = 0.2,
               size = .2) +
  coord_map()


# then put sites on it...
map_with_sites <- 
  map_with_sites +
  geom_point(data = site_coords,
             aes(x = long, 
                 y = lat), 
             color="black", 
             size = 3) +
  geom_point(data = site_coords,
             aes(x = long, 
                 y = lat), 
             color="white", 
             size = 2) 

# then put labels on it...
# get a halo on the text https://stackoverflow.com/a/10691826/1036500
theta <- seq(pi/8, 2*pi, length.out=16)
xo <- diff(range(site_coords$long))/200
yo <- diff(range(site_coords$lat))/200
for(i in theta) {
  map_with_sites <- 
    map_with_sites + geom_text_repel( data = site_coords,
                        aes_q(x = bquote(long + .(cos(i)*xo)),
                              y = bquote(lat + .(sin(i)*yo)),
                              label = ~label_number), 
                        size=3, 
                        alpha=1/2,
                        colour='white',
                        seed = 1, 
                        point.padding = .1)
}

# add text over halo for site names
map_with_sites <- 
  map_with_sites + 
  geom_text_repel(data = site_coords,
             aes(x = long, 
                 y = lat,
                 label = label_number), 
             size=3, 
             colour='black',
             seed = 1, 
             point.padding = .1) 

# add region labels, halos
for(i in theta) {
  map_with_sites <-
    map_with_sites +
    geom_text(
      data = region_label_coords,
      aes_q(
        x = bquote(long + .(cos(i) * xo)),
        y = bquote(lat + .(sin(i) *
                             yo)),
        label = ~label_name),
      size = 3,
      alpha = 1 / 2,
      colour = 'white')
}

# and region label core text
map_with_sites <- 
  map_with_sites + 
  geom_text(data = region_label_coords,
            aes(x = long,
                y = lat,
                label = label_name),
            size = 3,
            colour = "black")

# then add north arrow and scale
map_with_sites +
    scale_bar(lon = 110, 
              lat = -45,
              distance_lon = 500, 
              distance_lat = 100,
              distance_legend = 200, 
              dist_unit = "km")

ggsave("../figures/Fullagar-Hayes-Marwick-site-map.tiff",
       h = 7,
       w = 7)

```

# Analyse the bibliography

Here we take the MS Word document file with the bibliographic items and separate by report/thesis/paper, extract the year of publication, and plot the frequency of publications in each year for each category. 

```{r}
library(textreadr) # only for docx
library(tidyverse)

# data in ---------------------------------------------------------------
docx_doc <- "../data/raw_data/UWRDatabase.docx"

# read in the text
docx_doc_text <- 
  docx_doc %>%
  read_docx() 

# split into sections "THESES" "PUBLICATIONS"  and "UNPUBLISHED REPORTS" ------
splitAt <- function(x, pos) {pos <- c(1L, pos, length(x) + 1L); Map(function(x, i, j) x[i:j], list(x), head(pos, -1L), tail(pos, -1L) - 1L)}

docx_doc_text_split <- 
  splitAt(docx_doc_text, c(which(docx_doc_text == "PUBLICATIONS"),
                           which(docx_doc_text ==  "THESES"),
                           which(docx_doc_text == "UNPUBLISHED REPORTS")))

docx_doc_text_split <- docx_doc_text_split[-1]

# how many items in each section?
map_int(docx_doc_text_split, length) # 170, 50, 71

# Extract year of publication -------------------
# get dates of publication for each 

test <- c('1800, 1950 2010 3020 1930-1940 (1960), 1970. 1960peanuts, 1980a.' )

library(stringi)
get_dates <- function(x){
  strsplit(x, " ") %>% 
    unlist() %>% 
    stri_extract_first(., regex  = "(19|20)\\d{2}\\.|(19|20)\\d{2}[a-z]{1}\\.") %>% 
    stri_extract_first(., regex = "\\b(?<!\\-)[^\\s\\-]+\\b(?!-)") %>% 
    unlist() %>% 
    parse_number() %>% 
    as.numeric() %>% 
    discard(is.na(.))
}

# testing
get_dates(test)

# do it
years <- map(docx_doc_text_split, ~get_dates(.x))

# how many years did we get?
map_int(years, length) %>% discard(. == 0) # 169  49  72

# plot --------------------------------------------------------------------

names(years) <- c("Publications", "Theses", "Unpublished reports")
years_df <- unnest(bind_rows(enframe(years)))

years_df_summary <- 
  years_df %>% 
  group_by( name, value) %>% 
  tally()

# manual corrections

# publications

# 2010: 4
# 2017: 7

# reports
# 1994: 2
# 2010: 3

years_df_summary <- 
  years_df_summary %>% 
  mutate(n = case_when(
    name == "Publications" & value == 2010 ~ 4,
    name == "Publications" & value == 2017 ~ 7,
    name == "Unpublished reports" & value == 1994 ~ 2,
    name == "Unpublished reports" & value == 2010 ~ 3,
    TRUE ~ as.numeric(n)
    ))


# pretty facet titles
library(glue)
publication_caption <-
  glue("Publication (n = {years_df_summary %>%  
       filter(name == 'Publications') %>% 
       summarise(sum = sum(n)) %>% 
       pull(sum)})")
thesis_caption <-
  glue("Thesis (n = {years_df_summary %>%  
       filter(name == 'Theses') %>% 
       summarise(sum = sum(n)) %>% 
       pull(sum)})")
report_caption <-
  glue("Report (n = {years_df_summary %>%  
       filter(name == 'Unpublished reports') %>% 
       summarise(sum = sum(n)) %>% 
       pull(sum)})")


facet_names <- list(
  'Publications' = publication_caption,
  'Theses' = thesis_caption,
  'Unpublished reports' = report_caption
)

labeller <- function(variable, value){
  return(facet_names[value])
}


ggplot(years_df_summary,
       aes(value, 
           n)) +
  geom_col() +
  facet_wrap( ~ name, 
              ncol = 1, 
              labeller = as_labeller(labeller)) +
  theme_minimal(base_size = 14) +
  ylab("Number of publications/reports/theses") +
  xlab("") +
  scale_x_continuous(breaks = min(years_df_summary$value): max(years_df_summary$value),
                     labels =min(years_df_summary$value): max(years_df_summary$value)) +
  theme(axis.text.x = element_text(angle = 90,
                                   hjust = 0,
                                   vjust = 0.5,
                                   size = 8),
        panel.grid.minor.x = element_blank())

ggsave("../figures/fig_1_documents_per_year_per_document_type.tiff",
       h = 5,
       w = 7)



```


### Colophon

This report was generated on `r Sys.time()` using the following computational environment and dependencies: 

```{r colophon, cache = FALSE}
# which R packages and versions?
devtools::session_info()
```

The current Git commit details are:

```{r}
# what commit is this file at? You may need to change the path value
# if your Rmd is not in analysis/paper/
git2r::repository("../..")
```
